Waggle run distributions quantify collective foraging in honey bee colonies

### 2022-03-14 analysis/manuscript/paper.Rmd
updated changes made by Elli. Going to read through fresh in the morning and update before sending to Vicent for final edits.

---
title: "Forage site distributions quantify collective foraging in honey bee colonies"
author: "Joseph Palmer, Ash Samuelson, Elli Leadbeater and Vincent Jansen"
output:
  word_document
always_allow_html: true
bibliography: paper-refs.bib
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(showtext)
showtext_auto()

@@

Alternative title:

- Waggle run durations quantify collective foraging in honey bee colonies
- Waggle run duration distributions quantify collective foraging in honey bee colonies
- Distributions of waggle dance durations quantifies collective foraging in honey bee colonies

Or something like this.

@@

Honeybee foraging is a collective behaviour that is directed through the extraordinary waggle dance communication system, whereby colony foraging effort is allocated through a series of feedback loops [@Frisch1993; @Seeley1995]. Despite having been described as the most astounding example of non-primate communication [@Frisch1993], recent evidence suggests that the benefits of using dance information vary across environments and are sometimes even negligible [@Dornhaus2006; @Price2019]. Yet, because we cannot easily quantify collective foraging, pinpointing which environments promote collective foraging is challenging. Here, we present a mathematical model which quantifies collective foraging by inferring the proportion of 'scout' and 'recruit' dances on a hive's dance floor. By applying this methodology to waggle dance data from twenty hives, we show that dance information may be used less in environments where resources are sparse, potentially because colonies are required to send out more scouts to find food in these circumstances. Our methodology provides a tool to quantify waggle dance use in undisturbed hives at a much larger scale than hitherto has been possible and to identify those ecological conditions that may have driven the evolution of this remarkable collective behaviour.

@@

Despite such impressive complexity, in some situations, colonies that are prevented from communicating via dances achieve equal or even greater foraging success than their wild-type counterparts [@Sherman2002; @Dornhaus2006; @Gruter2013; @Price2019]. This indicates that using the dance does not always provide a foraging advantage and suggests bees may ignore spatial information in dances in favour of independent foraging. The prevalent explanation for this apparent redundancy portrays dance-based collective foraging as an evolutionary anachronism that was useful in ancient tropical environments but offers little benefit in modern temperate environments [@Dornhaus2004; @Donaldson-Matasci2012; @Price2015].

Given that dancing comes at a significant time cost [@Seeley1988], this raises the issue of why all temperate *Apis mellifera* colonies continue to dance. The dual questions of what foraging selection pressures drove the evolution of this uniquely complex collective behaviour, and why it continues to exist, are thus yet to be satisfactorily answered. Empirical studies have shown that honeybee colonies use dance information flexibly in response to their environment [@Sherman2002; @Dornhaus2004; @Dornhaus2006], suggesting that bees cease to rely on dance information if it provides little benefit [@Price2019]. Identifying those environments whereby modern honeybee colonies forage collectively rather than through individual search is a key step towards understanding dance evolution [@Price2015].

**Box 1**

[figure of waggle dance, maybe have text in like a box or something? Or just the text]

@@

Overall, evidence that colony-level collective foraging offers any benefit over individual search has proved surprisingly hard to come by. Whilst it is entirely possible the dance no longer provides a benefit in the modern world, a key barrier to assessing waggle dance use is that this requires distinguishing between foraging trips made by
scouting and recruited individuals. Scout foraging trips occur when a food source is located through individual search behaviour, whilst recruit foraging trips occur when a forage site is found by using the information provided in the dance of another bee [@Seeley1995]. Scouting and recruiting are not fixed behavioural categories, because individual bees can engage in both over the course of their foraging lifetime, and foragers can dance on return from any successful trip irrespective of whether they were recruited to the forage site or found it individually [@Beekman2007]. So far, scouts and recruits are distinguished by means of the foragers' individual search history, which requires following individual bees and is time and labour intensive [@Seeley1995].

Here, we will use a different approach and develop a model that describes the proportion of foraging trips in which bees act as scouts, and as recruits, within a colony, based on the distribution of foraging distances reported by waggle dancing, encoded as the duration of the run phase of the dance [@Frisch1993]. We will use this model to infer the proportion of scouts and recruits on the dance floor from dance data, and to demonstrate the principle, apply this methodology to dance data from a collection of hives in different environments.

@@

```{r, fig.cap="Figure 1. A. Outline of individual foraging. Foragers leave the hive on a random flight path (white lines) and continue until they encounter a resource (flowers in the figure). B. Outline of foraging with recruitment. Some foragers (scouts) continue to identify resources but others (recruits) sample available dances provided by scouts and other recruits. C. Log inverse frequency distribution of foraging distances reported by scouts. D. Log inverse frequency distribution of foraging distances reported by recruits."}
knitr::include_graphics("../results/figures/simulation.png")
```

Scouts and recruits create different foraging distance distributions on the dance-floor. These different distance distributions can be used to quantify the relative importance of collective foraging. Scouting foragers report a distribution of distances that reflects their probability of encountering a resource in the environment (see methods), whilst foragers that recruits report distances that reflect the distribution of the most profitable resource available (see methods). On the dance-floor, the distribution of distances is modified through the number of runs performed for a resource, which depends on the profitability of a resource, which honeybees achieve by measuring the energetic efficiency of a foraging trip through the ratio of energetic gain to energetic cost [@Seeley1994A]. By combining this profitability with the distributions identified in our simulations, we develop a novel method to describe the distribution of waggle runs reported on the dance-floors of real honeybee colonies as a superposition of scout and recruit distributions (Fig 2, see Methods).

@@

To evaluate the use of waggle dance information and individual search in honeybee colonies foraging in 'natural' landscapes, we analysed an existing dataset of 3378 waggle dance observations from 20 observation hives, recorded between April-September 2017, (previously described in [@Samuelson2021]). Hives were situated at different locations in South East England (see methods, figure 3A) and were visited every two weeks for a period of 24 weeks. On each visit, two hours of continuous waggle dance data was recorded by training a camcorder onto the dance floor. The footage of the dances was decoded manually ([@Couvillon2012; @Schurch2019] to extract waggle run durations. For each site we fit both the collective and individual models and used model selection to determine which provided the better explanation of the data, and (if the collective model provided a better fit) to quantify the relative use of social information through estimating the parameter p. We calculated the goodness-of-fit using a Kolmorgorov-Smirnov (KS) test to ascertain if the model provided a plausible explanation of the data [@Goldstein2004].

@@

```{r, fig.cap="Figure 3. A. Location of study hives in Southern England coloured by which model, collective (black circles) or individual (red triangles) provided the best explanation of the data as indicated by lowest AIC score. Sites STU and ZSL (C & D) are marked. Inset plot shows the location of the sites within the UK. B. Distribution of goodness of fit confidence values for each model fit to waggle run durations from each site. P value is derived from a bootstrapped two-sided KS test comparing the fitted model predictions to the empirical data. The red dashed line marks the significance threshold of 0.05. Values greater than 0.05 indicate no statistically significant difference is observed between the model and the data, indicating the model provides a good fit to the data. C. Model fits to waggle run durations collected from the ZSL hive along with histogram showing distribution. D. Model fits to waggle run durations collected from the STU hive along with histogram showing distribution. Model fits show the compliment cumulative frequency distribution, giving the probability of sampling a value greater than or equal to x."}
knitr::include_graphics("../results/figures/sites_model_plot.png")
```

Our results show that, whilst colony-level foraging is mostly comprised of a mixture of scout and recruit foraging trips, in some circumstances, colony foraging can be better described by individual foraging alone. This suggests that in some environments honey bees engage in scouting to find new food sites rather than follow dances. Previous empirical studies have shown that individual foragers flexibly alter their use of waggle dance information in response to unrewarding resources [@Gruter2011; @Price2019]. Evaluating how these individual decisions influence the collective, however, has historically been a challenge as it is effectively impossible to track an individuals foraging behaviour over a landscape. As our results show that individual foraging accurately describes colony foraging in three different sites, these findings support the idea that individual decision making can dominate colony foraging and demonstrates further evidence for flexible waggle dance use by honey bee colonies.

Our finding that foraging in some landscapes is more accurately described by scouting supports the hypothesis that the dance is used flexibly in response to environmental factors. As the collective model quantifies the contribution of recruitment to colony foraging, the fit of our model provides a new tool to evaluate how environmental factors influence waggle dance use, and thus how a colony uses waggle dance information under different environmental circumstances.

We estimated the proportion of scout and recruit foraging trips for each individual colony and investigated whether these estimates might correlate with land-use. Each site used in this study is situated in a landscape comprised of different proportions of land-use types. To evaluate how these different land-use types influenced waggle dance use, we first classified the different land-use types of the area surrounding each site to obtain a standardised land-use profile for the urban and agri-rural environments separately as many land-use types present in urban areas do not occur in agri-rural environments and vice versa (see methods). As many of the resulting land-use types covaried with each other, we performed a Partial Least Squares (PLS) analysis [@Carrascal2009] (see methods) to determine the principal components that represent combinations of land-use types which explained the most variation in the proportion of scout dances within agri-rural and urban environments. As we have a small sample size, this analysis was jackknifed to evaluate the robustness of our results to influential points (see methods, supplementary material section _._)

In the agri-rural environment one combination of land use types explained ~52% of the variation in the proportion of scouts (beta regression: $R^2 = 0.52$, $\phi$ = 3.7, p < 0.05, Fig. 4. A). This principal component correlates positively with arable land and negatively with non-agricultural unmanaged green space, built-up areas and water (Fig. 4. B). These land-use types maintain a significant correlation with PC1 over the jackknifed PLS (Fig. 4. B) indicating the results are robust. As arable land increases whilst unmanaged green space and built-up areas decrease, the proportion of individuals prioritising private information increases. Arable land dominates the agri-rural environment accounting for over 28% of land-coverage, whilst built-up areas represent around 17% of lands (Table 1). Non-agricultural unmanaged green space represent smaller amounts of land, accounting for less than 3% of the total land combined (Table 1).

@@

A similar effect is also observed in the urban environment, where a single PC explained ~59% of the variance in the proportion of scouts (beta regression: $R^2$ = 0.59, $\phi$ = 6.2, p < 0.05, Fig. 5. A). This PC correlates positively with parks, allotments and cemeteries, whilst correlating negatively with sparse residential lands and railways (Fig. 5. B). However, the jackknifed PLS identifies CAD as a highly influential site, the removal of which causes sparse residential land to have a positive correlation with PC1 (Supplementary Material _._). Visualising how waggle dance use is influenced by changes in these land-use types shows that when the landscape is dominated by sparse residential land, as in site HHS, waggle dance use dominates the colonies foraging strategy. However, this reduces as sparse residential land is replaced by parks, allotments, and cemeteries (Fig. 5). Sparse residential land dominates the urban environment accounting for over 34% of land-coverage, whilst parks allotments and cemeteries and railway corresponds to ~8% and ~1 respectively (Table 2).

@@

In agri-rural environments, land-use types which typically associate with resource scarcity correlate positively with a higher proportion of scouts, whilst those which correlate with a higher resource abundance correlate with a lower proportion of scouts. Arable land is typically nutritionally poor for honey bees [@Samuelson2021]. On the contrary, built up areas often have gardens with ample flowers available for pollinators [@Baldock2019]. Although non-agricultural unmanaged green space and water show as significantly correlating with reduced use of the dance, as they cover such a small amount of area, combined with a small sample size of only 10 sites, we cannot ignore the possibility that their significance is a random effect rather than evidence of a biological mechanism. Overall, these findings indicate that in agri-rural environments, when resources are more abundant and thus easier to find, the waggle dance is used more, presumably to identify the best resources of those on offer. Conversely, where resources are sparse and more difficult to find, a decreased use of the waggle dance suggests the colony focuses more on individual exploration rather than in trying to locate the best resources.

A similar result is observed in urban sites, where the presence of flower rich gardens in residential areas appears to promote increased use of the dance. Conversely, an increase in the amount of parks, allotments and cemeteries appears to decrease the use of the dance. Such land-use types could be expected to be quite resource rich for pollinators yet our findings suggests honey bee colonies decrease the use of the dance in such environments. It is possible that such environments vary seasonally in terms of their forage availability, allowing them to function both as floral deserts and oasis at different times of the year. As our waggle dance observations were collected over the course of an entire season, we may only be picking up the dominant signal over this time frame. However, the jackknife analysis shows our results are highly sensitive to influential points. Further data is therefore needed to investigate how the urban environment influences waggle dance use in honey bees.

The fitting of our model to waggle dance observations provides a novel tool to quantify the relative importance of social and private information in honeybees. Further, by identifying the land-use combinations which most influence variation in the proportion of scouts, our analysis allows us to map the environment along the major axis of honeybee information use and visualise how land-use influences the use of the waggle dance. Our results within the agri-rural sites are in broad agreement with other studies which evaluated foraging performance with and without the dance [@Dornhaus2006; Price2019]. These results suggest that, even though collective foraging is not always beneficial, that in environments where high quality resources are present, but relatively scarce, collective foraging is being carried out by colonies, suggesting that under these conditions exchanging social information through the waggle dance confers benefits. These benefits of dancing appear to outweigh the costs associated with dancing and could explain why this remarkable behaviour is present in modern-day bees. However, the results for our urban sites do not produce a clear biological rational and present an all together more varied picture. More waggle dance data from urban colonies is therefore required to further probe how honeybees use the waggle dance in this heavily anthropogenic environment.

>>>>

---
title: "Waggle run distributions quantify collective foraging in honey bee colonies"
author: "Joseph Palmer, Ash Samuelson, Elli Leadbeater and Vincent Jansen"
output:
  word_document
always_allow_html: true
bibliography: paper-refs.bib
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(showtext)
showtext_auto()

@@

@@

Honeybee foraging is an extraordinary collective behaviour that is directed through the celebrated waggle dance communication system, whereby colony foraging effort is allocated through a series of feedback loops [@Frisch1993; @Seeley1995]. Here, we present a mathematical model that quantifies the extent to which honeybee colonies behave collectively in different ecological environments by inferring the proportion of dances performed by bees that have explored individually, and those that have been recruited to food sources. By applying this methodology to waggle dance data from twenty hives, we show that colonies are less reliant on dance information where resources are sparse, potentially because colonies are required to send out more scouts to find food in these circumstances. Our methodology provides a tool to identify those ecological conditions in which honeybee colonies rely on dance communication, and so to understand the selection pressures that may have driven the evolution of this remarkable collective behaviour.

@@

Despite such impressive complexity, a wealth of research now shows that foraging bees frequently do not use dance information to find forage sites [@Gruter2009]. Accordingly, in many situations, colonies that are prevented from communicating via dances achieve equal or even greater foraging success than their wild-type counterparts [@Sherman2002; @Dornhaus2006; @Gruter2013; @Price2019]. This apparent variation in the use of collective behaviour is intriguing because it may provide clues as to those ecological selection pressures that were critical in the evolution of the dance communication system. Despite the extraordinary amount of research interest that has focused on the proximate mechanisms of collective decision-making via dance communication, the ultimate evolutionary drivers of its unique evolution within *Apis* -unparalleled in even a single other social insect- remain obscure [@Price2015]. Identifying those circumstances in which bees use the dance is likely to be a key step towards understanding when it is likely to afford fitness benefits at the colony level, but attempts to do so are challenging because they require constant monitoring of real-world colony weight [@Kirchner1998; @Sherman2002; @Dornhaus2006; @Gruter2013; @Nurnberger2017; @Price2019] -a noisy proxy of foraging success that is influenced by many other abiotic and biotic factors, including colony health- over ecologically credible time periods [@Schurch2014]. To do this at a scale that allows inter-colony variation in foraging environments at sufficient replication is a major logistical hurdle (but see [@Nurnberger2017]).

**Figure of honeybee foraging**

[figure of waggle dance]

@@

Here, we present a a model that can identify the proportion of foraging trips that are made by bees recruited through waggle dances, and thus quantify collective behaviour, in unmanipulated hives with no requirement to manipulate hive orientations or assay foraging efficiency. Our model seeks to distinguish between “scout” foraging trips, in which food sources that have been located through individual search are reported by waggle dancing, and “recruit” trips in which a forage site is found by using the information provided in the dance of another bee [@Seeley1995]. Scouting and recruiting are not fixed behavioural categories, because individual bees can engage in both over the course of their foraging lifetime, and foragers can dance on return from any successful trip irrespective of whether they were recruited to the forage site or found it individually [@Beekman2007]. In a hypothetical colony that relied only on individual search to find foraging sites, all trips would be “scout” trips, but as recruitment becomes more important, the proportion of “recruit” trips will increase. Previously, scout and recruit trips have only been distinguishable by means of the foragers’ individual search history, which requires following individual bees and is time and labour intensive [@Seeley1995].

@@

```{r, fig.cap="Figure 1. A. Outline of individual foraging. Foragers leave the hive on a random flight path (white lines) and continue until they encounter a resource (flowers in the figure). B. Outline of foraging with recruitment. Some foragers (scouts) continue to identify resources but others (recruits) sample available dances provided by scouts and other recruits. C. Log inverse frequency distribution of foraging distances reported by scouts. D. Log inverse frequency distribution of foraging distances reported by recruits. Here, inverse frequency distribution refers to probability of sampling a value greater than $X$."}
knitr::include_graphics("../results/figures/simulation.png")
```

Scouts and recruits create different waggle run distributions on the dance-floor. These different distributions can be used to quantify the relative importance of collective foraging. Scouting foragers report a distribution of distances that reflects their probability of encountering a resource in the environment (see methods), whilst recruits report distances that reflect the distribution of the most profitable resources available (see methods). On the dance-floor, the distribution of distances is modified through the number of runs performed for a resource, which depends on the profitability of a resource, which honeybees achieve by measuring the energetic efficiency of a foraging trip through the ratio of energetic gain to energetic cost [@Seeley1994A]. By combining this profitability with the distributions identified in our simulations, we develop a novel method to describe the distribution of waggle runs reported on the dance-floors of real honeybee colonies as a superposition of scout and recruit distributions (Fig 2, see Methods).

@@

To evaluate the use of waggle dance information and individual search in honeybee colonies foraging in 'natural' landscapes, we analysed a pre-existing dataset of 3378 waggle dance observations from 20 observation hives, recorded between April-September 2017, (previously described in [@Samuelson2021]). Hives that contributed to this dataset had been situated at different locations in South East England (see methods, figure 3A) and visited every two weeks for a period of 24 weeks. On each visit, two hours of continuous waggle dance data was recorded by training a camcorder onto the dance floor. The footage of the dances was decoded manually [@Couvillon2012; @Schurch2019] to extract waggle run durations. Using this data, for each site we fit both the collective and individual models and used model selection to determine which provided the better explanation of the data, and (if the collective model provided a better fit) to quantify the relative use of social information through estimating the parameter p. We calculated the goodness-of-fit using a Kolmorgorov-Smirnov (KS) test to ascertain if the model provided a plausible explanation of the data [@Goldstein2004].

@@

```{r, fig.cap="Figure 3. A. Location of study hives in Southern England coloured by which model, collective (black circles) or individual (red triangles) provided the best explanation of the data as indicated by lowest AIC score. Sites STU and ZSL (C & D) are marked. Inset plot shows the location of the sites within the UK; shaded area in the main plot indicates Greater London. B. Distribution of goodness of fit confidence values for each model fit to waggle run durations from each site. P value is derived from a bootstrapped two-sided KS test comparing the fitted model predictions to the empirical data. The red dashed line marks the significance threshold of 0.05. Values greater than 0.05 indicate no statistically significant difference is observed between the model and the data, indicating the model provides a good fit to the data. C. Model fits to waggle run durations collected from the ZSL hive, for which the model did not outperform the individual model, along with histogram showing distribution. D. Model fits to waggle run durations collected from the STU hive along with histogram showing distribution. Model fits show the compliment cumulative frequency distribution, giving the probability of sampling a value greater than or equal to x."}
knitr::include_graphics("../results/figures/sites_model_plot.png")
```

These results show that, whilst colony-level foraging is mostly comprised of a mixture of scout and recruit foraging trips, in some circumstances, colony foraging can be better described by individual foraging alone. This suggests that in some environments honey bees mainly engage in scouting to find new food sites rather than follow dances. Previous empirical studies have shown that individual foragers may flexibly alter their use of waggle dance information if its use is proving unproductive [@Gruter2011; @Price2019]. Evaluating how these individual decisions influence the collective, however, has historically been a challenge as it is effectively impossible to track an individual's foraging behaviour over a landscape. As our results show that individual foraging accurately describes colony foraging in three different sites, these findings support the idea that individual decision making can dominate colony foraging and demonstrates further evidence for flexible waggle dance use by honey bee colonies.

Further quantification of the use of waggle-dance recruitment within all colonies, as a proportion of all foraging trips, can be achieved by extracting the estimated proportion of scout trips (p) for each site. Since our sites varied in land-use characteristics and potentially thus forage distributions (although in this case, not by design for this study), we investigated whether these estimates might correlate with land-use. We first classified the different land-use types of the area surrounding each site [@Samuelson2018] to obtain a standardised land-use profile for the urban and agri-rural environments separately as many land-use types present in urban areas do not occur in agri-rural environments and vice versa (see methods). We then performed a Partial Least Squares (PLS) analysis [@Carrascal2009] (see methods) to determine the principal components that represent combinations of land-use types which explained the most variation in the proportion of scout dances within agri-rural and urban environments. As we have a small sample size (20 sites; 10 urban and 10 agri-rural), this analysis was jackknifed to evaluate the robustness of our results to influential points (see methods, supplementary material section _._)

In the agri-rural environments one combination of land use types explained ~52% of the variation in the proportion of scouts (beta regression: $R^2 = 0.52$, $\phi$ = 3.7, p < 0.05, Fig. 4. A). This principal component correlates positively with arable land (29% of land coverage; Table 1) and negatively with built-up areas (17% of land coverage) ; it also correlates negatively with non-agricultrual unmanaged green space and water , but together these represent less than 3% of land-use (Fig. 4. B). These land-use types maintain a significant correlation with PC1 over the jackknifed PLS (Fig. 4. B) indicating the results are robust. As arable land increases whilst built-up areas decrease, the proportion of trips that are driven by individual search increases.

@@

A similar effect was observed in the urban environment, where a single PC explained ~59% of the variance in the proportion of scouts (beta regression: $R^2$ = 0.59, $\phi$ = 6.2, p < 0.05, Fig. 5. A). This PC correlates positively with parks, allotments and cemeteries (8% of land cover; Table 2), whilst correlating negatively with sparse residential lands (35% of land cover) and railways but these represent only 1% of land cover. (Fig. 5. B). However, we have low confidence in this finding because the jackknifed PLS identifies CAD as a highly influential site, the removal of which causes sparse residential land to have a positive correlation with PC1 (Supplementary Material _._).

@@

Across both environments our results hint at a decrease in reliance on waggle dance recruitment as resources become harder to find and foraging trips become longer [@Samuelson2021]. Arable land is likely to be nutritionally poor for pollinators (note that oilseed rape field and fruit crops were not included in this category), while residential areas -particularly gardens- are considered nutritional hotspots for bees [@Baldock2019]. Similarly, in the urban environments sparse-residential coverage correlated negatively with the proportion of scout trips. However, we are cautious in this interpretation for two reasons. Firstly, because our study was designed as a proof-of-principle based on an existing dataset, we did not systematically assess the availability of forage within these sites. Secondly, although we saw a similar pattern in the urban data for sparse residential areas (35% of land-cover), which correlated negatively with the proportion of scout trips, this pattern reversed when a highly influential site (CAD; identified by the jacknifed PLS) was removed, and we also found parks, allotments and cemeteries (8% land cover), which could be expected to provide abundant forage for bees [@Baldock2019], to correlate positively with independent search.

The fitting of our model to waggle dance observations provides a novel tool to quantify the relative importance of social and private information in honeybees. Whilst recruitment is well known to occur through other mechanisms besides the waggle dance, such as through olfactory cues, the exceptionally close fit of our model to waggle run durations underlines the importance of the waggle dance in honey bee foraging. Further, by identifying the land-use combinations which most influence variation in the proportion of scouts, our analysis allows us to map the environment along the major axis of honeybee information use and visualise how land-use influences the use of the waggle dance.

Our results within the agri-rural sites are in broad agreement with other studies which evaluated foraging performance with and without the dance [@Dornhaus2006; Price2019]. These results suggest that, even though collective foraging is not always beneficial, that in environments where high quality resources are present, but relatively scarce, collective foraging is being carried out by colonies, suggesting that under these conditions exchanging social information through the waggle dance confers benefits. These benefits of dancing appear to outweigh the costs associated with dancing and could explain why this remarkable behaviour is present in modern-day bees. However, the results for our urban sites do not produce a clear biological rational and present an all together more varied picture. An alternative hypothesis could be that the the dance is influenced by resource stability more than abundance and that longer term effects may contribute to the benefits of the dance [@Schurch2014]. Whilst the limiting number of sites used in this study handicaps our ability to progress this debate here, as a proof of concept, the methods derived here provide a new toolkit for evaluating factors influencing waggle dance recruitment
